### 2023-03-29 response-to-audit-report.md

 -- | Check that parameters in the datum are not changed by the transaction.
-checkParamsOutput :: Property
-checkParamsOutput =
-  checkDatumOutput
+checkParamsOutput :: [ReferencePath] -> Property
+checkParamsOutput referencePaths =
+  checkDatumOutput referencePaths
     $ \marloweData ->
       do
         -- Replace the output parameters with a random one.
@@ -448,9 +565,9 @@ checkParamsOutput =

 -- | Check that state output to a script matches its semantic output.
-checkStateOutput :: Property
-checkStateOutput =
-  checkDatumOutput
+checkStateOutput :: [ReferencePath] -> Property
+checkStateOutput referencePaths =
+  checkDatumOutput referencePaths
     $ \marloweData ->
       do
         -- Replace the output state with a random one.
@@ -460,9 +577,9 @@ checkStateOutput =

 -- | Check that contract output to a script matches its semantic output.
-checkContractOutput :: Property
-checkContractOutput =
-  checkDatumOutput
+checkContractOutput :: [ReferencePath] -> Property
+checkContractOutput referencePaths =
+  checkDatumOutput referencePaths
     $ \marloweData ->
       do
         -- Replace the output ccontact with a random one.
@@ -482,8 +599,8 @@ hasMerkleizedInput =

 -- | Check than an invalid merkleization is rejected.
-checkMerkleization :: Bool -> Property
-checkMerkleization valid =
+checkMerkleization :: [ReferencePath] -> Bool -> Property
+checkMerkleization referencePaths valid =
   let
     -- Merkleizedd the contract and its input.
     modifyBefore = merkleize
@@ -499,12 +616,12 @@ checkMerkleization valid =
                hashes <- input `uses` (concatMap merkleHash . txInputs)
                infoData %= (AM.fromList . filter ((`notElem` hashes) . fst) . AM.toList)
   in
-    checkSemanticsTransaction modifyBefore modifyAfter hasMerkleizedInput valid False
+    checkSemanticsTransaction referencePaths modifyBefore modifyAfter hasMerkleizedInput valid False False

 -- | Check that non-positive accounts are rejected.
-checkPositiveAccounts :: Property
-checkPositiveAccounts =
+checkPositiveAccounts :: [ReferencePath] -> Property
+checkPositiveAccounts referencePaths =
   let
     modifyBefore =
       do
@@ -515,7 +632,7 @@ checkPositiveAccounts =
         -- Add the non-positive entry to the accounts.
         inputState %= (\state -> state {accounts = AM.insert (account, token) amount' $ accounts state})
   in
-    checkSemanticsTransaction modifyBefore noModify noVeto False False
+    checkSemanticsTransaction referencePaths modifyBefore noModify noVeto False False False

 -- | Compute the authorization for an input.
@@ -547,8 +664,8 @@ deleteFirst f z =

 -- | Check that a missing authorization causes failure.
-checkAuthorization :: Property
-checkAuthorization =
+checkAuthorization :: [ReferencePath] -> Property
+checkAuthorization referencePaths =
   let
     modifyAfter =
       do
@@ -565,7 +682,7 @@ checkAuthorization =
         -- Remove the first PKH signatory.
         infoSignatories %= deleteFirst matchPkh
   in
-    checkSemanticsTransaction noModify modifyAfter hasAuthorizations False False
+    checkSemanticsTransaction referencePaths noModify modifyAfter hasAuthorizations False False False

 -- | Determine whether there are any external payments in a transaction.
@@ -585,8 +702,8 @@ decrementValue = foldMap (\(c, n, i) -> singleton c n (i - 1)) . flattenValue

 -- | Check that an insufficient payment causes failure.
-checkPayment :: Property
-checkPayment =
+checkPayment :: [ReferencePath] -> Property
+checkPayment referencePaths =
   let
     modifyAfter =
       do
@@ -600,7 +717,7 @@ checkPayment =
         -- Update the outputs.
         infoOutputs %= fmap decrementPayment
   in
-    checkSemanticsTransaction noModify modifyAfter hasExternalPayments False False
+    checkSemanticsTransaction referencePaths noModify modifyAfter hasExternalPayments False False False

 -- | Remove a role input UTxOs from the transaction.
diff --git a/marlowe-test/src/Spec/Marlowe/Plutus/Transaction.hs b/marlowe-test/src/Spec/Marlowe/Plutus/Transaction.hs
index 9141d071f..88201e361 100644
--- a/marlowe-test/src/Spec/Marlowe/Plutus/Transaction.hs
+++ b/marlowe-test/src/Spec/Marlowe/Plutus/Transaction.hs
@@ -27,6 +27,7 @@ module Spec.Marlowe.Plutus.Transaction
   , noModify
   , shuffle
     -- * Conditions
+  , isScriptTxIn
   , noVeto
   ) where

@@ -61,7 +62,7 @@ import Plutus.Script.Utils.Scripts (datumHash)
 import Plutus.V1.Ledger.Value (gt)
 import Plutus.V2.Ledger.Api
   ( Address(Address)
-  , Credential(PubKeyCredential)
+  , Credential(..)
   , CurrencySymbol
   , Datum(..)
   , DatumHash(..)
@@ -76,7 +77,7 @@ import Plutus.V2.Ledger.Api
   , ToData(toBuiltinData)
   , TxInInfo(..)
   , TxInfo(..)
-  , TxOut(TxOut)
+  , TxOut(TxOut, txOutAddress)
   , UpperBound(UpperBound)
   , Value
   )
@@ -105,10 +106,11 @@ import Spec.Marlowe.Plutus.Types
   , role
   , scriptPurpose
   )
+import Spec.Marlowe.Reference (ReferencePath, arbitraryReferenceTransaction)
 import Spec.Marlowe.Semantics.Arbitrary (arbitraryGoldenTransaction, arbitraryPositiveInteger)
 import Spec.Marlowe.Semantics.Golden (GoldenTransaction)
 import Spec.Marlowe.Semantics.Merkle (deepMerkleize, merkleizeInputs)
-import Test.Tasty.QuickCheck (Arbitrary(..), Gen, elements, suchThat)
+import Test.Tasty.QuickCheck (Arbitrary(..), Gen, elements, frequency, listOf, suchThat)

 import qualified Language.Marlowe.Core.V1.Semantics.Types as M (Party(Address))
 import qualified Plutus.V1.Ledger.Value as V (adaSymbol, adaToken, singleton)
@@ -185,11 +187,15 @@ makeDeposit :: Input
 makeDeposit input' =
   do
     ref <- lift arbitrary
-    address' <- lift arbitrary
+    address' <- lift $ arbitrary `suchThat` notScriptAddress
     pure
       $ case getInputContent input' of
-          IDeposit _ (M.Address _ address) (Token c n) i -> pure . TxInInfo ref $ TxOut address  (V.singleton c n i) NoOutputDatum  Nothing
-          IDeposit _ (Role _             ) (Token c n) i -> pure . TxInInfo ref $ TxOut address' (V.singleton c n i) NoOutputDatum  Nothing
+          IDeposit _ (M.Address _ address) (Token c n) i -> if i > 0
+                                                              then pure . TxInInfo ref $ TxOut address  (V.singleton c n i) NoOutputDatum  Nothing
+                                                              else mempty
+          IDeposit _ (Role _             ) (Token c n) i -> if i > 0
+                                                              then pure . TxInInfo ref $ TxOut address' (V.singleton c n i) NoOutputDatum  Nothing
+                                                              else mempty
           _                                              -> mempty

 -- | Create role input for a Marlowe semantics transaction.
@@ -199,7 +205,7 @@ makeRoleIn input' =
   do
     MarloweParams currencySymbol <- use marloweParams
     ref <- lift arbitrary
-    address  <- lift arbitrary
+    address  <- lift $ arbitrary `suchThat` notScriptAddress
     pure
       $ case getInputContent input' of
           IDeposit _ (Role role') _ _         -> pure . TxInInfo ref $ TxOut address (V.singleton currencySymbol role' 1) NoOutputDatum Nothing
@@ -276,6 +282,21 @@ makeSpendSignatory (TxInInfo _ (TxOut (Address (PubKeyCredential pkh) _) _ _ _))
 makeSpendSignatory _                                                             = mempty

+-- | Combine payments with the same address and hash.
+consolidatePayments :: [TxOut] -> [TxOut]
+consolidatePayments ps =
+  let
+    extractAddressHash (TxOut address _ hash _) = (address, hash)
+    extractValue (TxOut _ value _ _) = value
+  in
+    [
+      TxOut address value hash Nothing
+    |
+      ah@(address, hash) <- nub $ extractAddressHash <$> ps
+    , let value = foldMap extractValue $ filter ((== ah) . extractAddressHash) ps
+    ]
+
+
 -- | Generate a valid Marlowe semantics transaction.
 validSemanticsTransaction :: Bool                                          -- ^ Whether to add noise to the script context.
                           -> ArbitraryTransaction SemanticsTransaction ()  -- ^ The generator.
@@ -313,7 +334,7 @@ validSemanticsTransaction noisy =
     MarloweParams currencySymbol <- use marloweParams
     -- Add the payments.
     (payments, paymentData) <- fmap (bimap mconcat mconcat . unzip) . mapM (makePayment currencySymbol) =<< (output `uses` txOutPayments)
-    infoOutputs <>= payments
+    infoOutputs <>= consolidatePayments payments
     infoData <>= AM.fromList paymentData

     -- Add the signatories.
@@ -336,13 +357,20 @@ validSemanticsTransaction noisy =

 -- | Generate an arbitrary, valid Marlowe semantics transaction: datum, redeemer, and script context.
-arbitrarySemanticsTransaction :: ArbitraryTransaction SemanticsTransaction ()  -- ^ Modifications to make before building the valid transaction.
+arbitrarySemanticsTransaction :: [ReferencePath]                               -- ^ The reference execution paths from which to choose.
+                              -> ArbitraryTransaction SemanticsTransaction ()  -- ^ Modifications to make before building the valid transaction.
                               -> ArbitraryTransaction SemanticsTransaction ()  -- ^ Modifications to make after building the valid transaction.
                               -> Bool                                          -- ^ Whether to add noise to the script context.
+                              -> Bool                                          -- ^ Whether to allow merkleization.
                               -> Gen (PlutusTransaction SemanticsTransaction)  -- ^ The generator.
-arbitrarySemanticsTransaction modifyBefore modifyAfter noisy =
+arbitrarySemanticsTransaction referencePaths modifyBefore modifyAfter noisy allowMerkleization =
   do
-    golden <- arbitraryGoldenTransaction
+    golden <-
+      frequency
+        [
+          (1, arbitraryGoldenTransaction allowMerkleization)  -- Manually vetted transactions.
+        , (5, arbitraryReferenceTransaction referencePaths)   -- Transactions generated using `getAllInputs` and `computeTransaction`.
+        ]
     start <- bareSemanticsTransaction golden
     (modifyBefore >> validSemanticsTransaction noisy >> modifyAfter)
       `execStateT` start
@@ -377,7 +405,7 @@ makePayoutRoleIn :: ArbitraryTransaction PayoutTransaction TxInInfo
 makePayoutRoleIn =
   do
     ref <- lift arbitrary
-    address <- lift arbitrary
+    address <- lift $ arbitrary `suchThat` notScriptAddress
     value <- V.singleton <$> marloweParamsPayout `uses` rolesCurrency <*> use role <*> pure 1
     pure
       . TxInInfo ref
@@ -445,15 +473,50 @@ validPayoutTransaction noisy =
     infoFee <~ V.singleton V.adaSymbol V.adaToken <$> lift arbitraryPositiveInteger

     -- Add noise.
-    when noisy addNoise
+    when noisy addNoise'

     -- Shuffle.
     shuffle

+-- | Check that an address is not for a script.
+notScriptAddress :: Address -> Bool
+notScriptAddress (Address (ScriptCredential _) _) = False
+notScriptAddress _ = True
+
+
+-- | Check that an input is not from a script.
+notScriptTxIn :: TxInInfo -> Bool
+notScriptTxIn = not . isScriptTxIn
+
+
+-- | Check that an input is from a script.
+isScriptTxIn :: TxInInfo -> Bool
+isScriptTxIn TxInInfo{txInInfoResolved=TxOut{txOutAddress=Address (ScriptCredential _) _}} = True
+isScriptTxIn _ = False
+
+
 -- | Add noise to the inputs, outputs, and data in a Plutus transaction.
-addNoise :: ArbitraryTransaction a ()
+addNoise :: ArbitraryTransaction SemanticsTransaction ()
 addNoise =
+  do
+    let
+      isPayout (Payment _ (Party _) _ i) = i > 0
+      isPayout _ = False
+    hasPayments <- any isPayout . txOutPayments <$> use output
+    let
+      arbitraryInput =
+        if hasPayments
+          then arbitrary `suchThat` notScriptTxIn
+          else arbitrary
+    infoInputs  <><~ lift (listOf arbitraryInput `suchThat` ((< 5) . length))
+    infoOutputs <><~ lift (arbitrary `suchThat` ((< 5) . length))
+    infoData    <><~ lift (fmap AM.fromList $ arbitrary `suchThat` ((< 5) . length))
+
+
+-- | Add noise to the inputs, outputs, and data in a Plutus transaction.
+addNoise' :: ArbitraryTransaction PayoutTransaction ()
+addNoise' =
   do
     infoInputs  <><~ lift (arbitrary `suchThat` ((< 5) . length))
     infoOutputs <><~ lift (arbitrary `suchThat` ((< 5) . length))
diff --git a/marlowe-test/src/Spec/Marlowe/Reference.hs b/marlowe-test/src/Spec/Marlowe/Reference.hs
new file mode 100644
index 000000000..4c3291e6a
--- /dev/null
+++ b/marlowe-test/src/Spec/Marlowe/Reference.hs
@@ -0,0 +1,156 @@
+-----------------------------------------------------------------------------
+--
+-- Module      :  $Headers
+-- License     :  Apache 2.0
+--
+-- Stability   :  Experimental
+-- Portability :  Portable
+--
+-- | Function to generate all valid transactions for contracts in JSON files.
+--
+-----------------------------------------------------------------------------
+
+
+{-# LANGUAGE DeriveGeneric #-}
+{-# LANGUAGE FlexibleContexts #-}
+{-# LANGUAGE LambdaCase #-}
+{-# LANGUAGE NumericUnderscores #-}
+{-# LANGUAGE OverloadedStrings #-}
+{-# LANGUAGE RecordWildCards #-}
+
+
+module Spec.Marlowe.Reference
+  ( -- * Types
+    ReferencePath(..)
+  , ReferenceTransaction(..)
+    -- * Analysis
+  , processContract
+    -- * Testing
+  , arbitraryReferenceTransaction
+  , readReferencePaths
+  , referenceFolder
+  ) where
+
+
+import Control.Monad (forM)
+import Control.Monad.Except (ExceptT(..), lift, throwError)
+import Data.Aeson (FromJSON, ToJSON, eitherDecodeFileStrict, encodeFile)
+import Data.Bifunctor (first)
+import Data.List (isSuffixOf)
+import GHC.Generics (Generic)
+import Language.Marlowe.Core.V1.Semantics (TransactionInput, TransactionOutput(..), computeTransaction)
+import Language.Marlowe.Core.V1.Semantics.Types (Contract, Party(Role), State(..), Token(..))
+import Language.Marlowe.FindInputs (getAllInputs)
+import Plutus.V2.Ledger.Api (POSIXTime)
+import Spec.Marlowe.Semantics.Golden (GoldenTransaction)
+import System.Directory (listDirectory)
+import System.FilePath ((</>))
+import Test.Tasty.QuickCheck (Gen, elements)
+
+import qualified PlutusTx.AssocMap as AM (empty, singleton)
+
+
+referenceFolder :: FilePath
+referenceFolder = "reference" </> "data"
+
+
+readReferencePaths :: IO [ReferencePath]
+readReferencePaths =
+  do
+    pathFiles <- fmap (referenceFolder </>) . filter (".paths" `isSuffixOf`) <$> listDirectory referenceFolder
+    fmap concat
+      . forM pathFiles
+      $ \pathFile ->
+        eitherDecodeFileStrict pathFile
+          >>= \case
+            Right paths -> pure $ filter (not . null . transactions) paths
+            Left msg -> error $ "Failed parsing " <> pathFile <> ": " <> msg <> "."
+
+
+arbitraryReferenceTransaction :: [ReferencePath] -> Gen GoldenTransaction
+arbitraryReferenceTransaction paths =
+  do
+    ReferencePath{..} <- elements paths
+    if length transactions > 1
+      then do
+             (ReferenceTransaction _ prior, ReferenceTransaction{..}) <- elements $ zip (init transactions) (tail transactions)
+             pure (txOutState prior, txOutContract prior, input, output)
+      else let
+             ReferenceTransaction{..} = head transactions
+            in
+              pure (state, contract, input, output)
+
+
+data ReferencePath =
+  ReferencePath
+  {
+    contract :: Contract
+  , state :: State
+  , transactions :: [ReferenceTransaction]
+  }
+    deriving (Generic, Show)
+
+instance FromJSON ReferencePath
+
+instance ToJSON ReferencePath
+
+
+data ReferenceTransaction =
+  ReferenceTransaction
+  {
+    input :: TransactionInput
+  , output :: TransactionOutput
+  }
+    deriving (Generic, Show)
+
+instance FromJSON ReferenceTransaction
+
+instance ToJSON ReferenceTransaction
+
+
+processContract
+  :: FilePath
+  -> FilePath
+  -> ExceptT String IO ()
+processContract contractFile pathsFile =
+  do
+    contract <- ExceptT $ first show <$> eitherDecodeFileStrict contractFile
+    traces <- ExceptT $ first show <$> getAllInputs contract
+    paths <- runTransactions contract `mapM` traces
+    lift $ encodeFile pathsFile paths
+
+
+runTransactions
+  :: Contract
+  -> (POSIXTime, [TransactionInput])
+  -> ExceptT String IO ReferencePath
+runTransactions contract (startTime, inputs) =
+  do
+    let
+      state = makeState startTime
+    transactions <- runTransaction contract state inputs
+    pure ReferencePath{..}
+
+
+runTransaction
+  :: Contract
+  -> State
+  -> [TransactionInput]
+  -> ExceptT String IO [ReferenceTransaction]
+runTransaction _ _ [] = pure []
+runTransaction contract state (input : inputs) =
+  case computeTransaction input state contract of
+    Error err -> throwError $ show err
+    output@TransactionOutput{..} -> (ReferenceTransaction{..} :) <$> runTransaction txOutContract txOutState inputs
+
+
+makeState
